This is the title [Article v\versionnumber]

### 2022-10-28 manuscript/manuscript.tex
Fix left-over merging conflict

Solvation thermodynamics govern any process involving solutes in a solvent.
%Solutes are constantly interacting with solvents in their environment, necessitating a good description of these interactions for accurate computational predictions. 
Especially the interaction with water is of utmost importance, as many biological processes occur in a aqueous environment. 
The free energy of solvation thereby strongly influences real-world problems such as hydrophobic effects, the binding of ligands to a biomolecule, as well as the dynamics and folding of proteins.

@@

For the visualization of three-dimensional contributions we require molecular visualization software that can handle \software{OpenDX} files.
We recommend using a current version of \software{PyMol}\cite{pymol} or \software{VMD}\cite{vmd}.
\section{Theory}

GIST is an implementation of inhomogeneous fluid solvation theory (IST) \cite{Lazaridis1998} that discretizes the free energy of solvation $\dgsolv$ on a three-dimensional grid. 
It was first devised by Nguyen et al. \cite{Nguyen2012} and its implementation in \software{cpptraj} was thoroughly described by Ramsey et al. \cite{Ramsey2016}. % to overcome the limitation of IST which was commonly used for regions of high water occupancy and its implementation in \software{cpptraj} was thoroughly described in \cite{Ramsey2016}.
Here, we only provide a short overview of the theory behind GIST.
For more detailed information, we recommend one of the more recent publications on developments in GIST \cite{Kraml2020,Chen2021}.

\section{Running GIST}
%\subsection{Running the GIST analysis}
To run a gist analyis, the following is required:
\begin{itemize}
	\item A running installation of \software{cpptraj}.
	\item A topology of the molecule of interest in a solvent box.
	\item An MD trajectory of that system.
\end{itemize}
GIST analysis can be divided in three steps: defining a region of interest, running GIST using \software{cpptraj}, and post-processing and analysis.
\cite{Ramsey2016}

\paragraph{Identify and define the region of interest}
In order to run a GIST analysis, the region where the solvent properties will be calculated must be first defined. This region is represented by a rectangular prism centered at (x, y, z) with $N_x \times N_y \times N_z$ voxels, where each voxel is cubic with a sidelength of \inlinecode{gridspacn}.
While integrals over GIST quantities are independent of the voxel size, individual voxel values converge slower at high resolutions.\pdfcomment{The calculation actually can actually get faster at higher resolution, because the same grid is used in the nearest-neighbor search.}
A default value of 0.5 A is used that provides reasonable discretization for many use-cases.

In \software{ccptraj}, the rectangular prism is defined by parameters of the \inlinecode{gist} command.
The center of the grid is defined by \inlinecode{gridcntr}, the dimensions of 
the grid is defined by \inlinecode{griddim} (in voxel counts), and the spacing 
is defined by \inlinecode{gridspacn} (in Å).

@@

The region can be defined using a ligand in a specific binding pocket of the 
protein, a specific residue of the protein, or the whole protein itself. A 
helper script FindCentroid.py can be used to determine the centroid of the 
region of interest as well as the dimensions of the rectangular prism that will 
fully enclose the region. A pdb file is needed as an input and the rectangular 
prism will be centered based on the centroid of the given structure.

@@

\paragraph{Run GIST using cpptraj}
GIST is incorporated in cpptraj which requires the topology and the trajectory of the MD simulation to be analayzed. The commands needed to run GIST in cpptraj can be written in an input file (e.g gist.in) for convenience. The basic comands, assuming a given grid center and a grid size of $30\times30\times30$ voxels with \SI{0.5}{\angstrom} spacing, are:

@@

The first line specifies the amber topology file that was used to run MD. The second line species the trajectory file and what frames to be analyzed. The third line tells cpptraj to run a GIST calculation using the region defined and print the output of the calculation in a file called gist.dat. The last instructs cpptraj to run the calculation. The syntax for running cpptraj given this input file is:

@@

\paragraph{Postprocessing and Analyzing GIST results}
Cpptraj outputs the quantities calculated as a "open data explorer" (.dx) file format which can be easily visualized in Pymol or VMD.
Aside from the thermodynamic quantities, GIST also calculates the number densities of hydrogen and oxygen in each voxel. Here is a full list of quantities calculated by GIST \todo{This is incomplete!}:

\begin{itemize}
	\item{[gX] For every element in the main solvent, the number density of atoms found in the voxel, in units of the bulk density. If the same element occurs multiple times, the bulk density is scaled accordingly. For water, gO and gH are produced.}
	%\item{[gH] Number density of hydrogen centers found in the voxel in units of the reference bulk density.}
	\item{[Esw] Mean solute-water interaction energy density.}
	\item{[Eww] Mean water-water interaction energy density.}
	\item{[PME] (only if PME was used) water PME energy density.}
	\item{[dTStrans] First order translational entropy density.}
	\item{[dTSorient] First order orientational entropy density .}
	\item{[neighbor] Mean number of waters neighboring the water molecules found in this voxel multiplied by the voxel number density.}
	\item{[dipole] Magnitude of mean dipole moment (polarization).}
	\item{[order] (if \inlinecode{doorder} was specified) Average Tetrahedral Order Parameter.}
	\item{[dipolex] x-component of the mean water dipole moment density}
	\item{[dipoley] y-component of the mean water dipole moment density}
	\item{[dipolez] z-component of the mean water dipole moment density}
	\item{[Eij] (if \inlinecode{doeij} was specified) Water-water interaction matrix.}
\end{itemize}

Several dx files are written by default, such as the energies, entropies, number density of oxygen, and number density of hydrogen. All quantities mentioned above are written to the output file (output.dat) and can be used to generate DX files for other quantities.

\section{Tutorial}
Here, we aim to guide the reader through a GIST analysis of the streptavidin-biotin complex.
In the process, the reader will learn how to set up and run a GIST calculation.
During the tutorial, we will
\begin{enumerate}
	\item compute the binding free energy of the complex from the individual energy contributions combined with the contribution of hydration computed using GIST.
	\item visualize local contributions to $\dgsolv$ in the streptavidin binding pocket as well as around biotin.
\end{enumerate}

@@

\subsection{Tutorial data}
We will use the 1STP crystal structure to start our calculations.
For the sake of simplicity, you can download prepared and solvated Amber topologies and structures based on 1STP by running
\begin{lstlisting}[style=bash]
git clone git@github.com:liedllab/gist-tutorial.git
\end{lstlisting}
This will download the tutorial files (including both the manuscript and the examples in the \software{code} folder).
\subsection{System Preparation and Equilibration}
If you want to skip this section, run 
\begin{lstlisting}[style=bash]
make equilibration-targets
\end{lstlisting}
in the code folder. This assumes that you have Python and Amber set up properly and that \software{pmemd.cuda} is in the PATH.

@@

@@

\begin{checklist}{The CPPTRAJ action}

>>>>

Solvation thermodynamics govern any process involving solutes in a solvent.
%Solutes are constantly interacting with solvents in their environment, necessitating a good description of these interactions for accurate computational predictions. 
The free energy of solvation is defined as the work needed to insert a solute into a solvent. Especially the interaction with water is of utmost importance, as many biological processes occur in a aqueous environment. 
It thereby strongly influences real-world problems such as hydrophobic effects, the binding of ligands to a biomolecule, as well as the dynamics and folding of proteins.

@@

For the visualization of three-dimensional contributions we require molecular visualization software that can handle \software{OpenDX} files.
We recommend using a current version of \software{PyMol}\cite{pymol} or \software{VMD}\cite{vmd}.

\section{Running GIST}
%\subsection{Running the GIST analysis}
To run a GIST analyis, the following is required:
\begin{itemize}
	\item A running installation of \software{cpptraj}.
	\item A topology of the molecule of interest in a solvent box.
	\item An MD trajectory of that system.
\end{itemize}
GIST analysis can be divided in three steps: defining a region of interest, running GIST using \software{cpptraj}, and post-processing and analysis.
\cite{Ramsey2016}

\paragraph{Identify and define the region of interest}
In order to run a GIST analysis, the region where the solvent properties will be calculated must be first defined. This region is represented by a rectangular grid centered at (x, y, z) with $N_x \times N_y \times N_z$ voxels, where each voxel is cubic with a sidelength of \inlinecode{gridspacn}.
While integrals over GIST quantities are independent of the voxel size, individual voxel values converge slower at high resolutions.\pdfcomment{The calculation actually can actually get faster at higher resolution, because the same grid is used in the nearest-neighbor search.}
The default value of \SI{0.5}{\angstrom} provides reasonable discretization for many use-cases.

The grid is defined by parameters of the \inlinecode{gist} command in \software{cpptraj}.
The grid center is defined by \inlinecode{gridcntr}, the voxel number in x, y, and z direction by \inlinecode{griddim}, and the spacing in \AA by \inlinecode{gridspacn}.

@@

Often, the region of interest will be defined by the surrounding of a molecule, which might either be a ligand in the binding pocket or also the whole protein. 
A helper script called \inlinecode{FindCentroid.py} for this purpose is supplied with this tutorial, and can be used as follows:

@@

\paragraph{Run GIST using cpptraj}
GIST is implemented in \software{cpptraj} which requires the topology and the trajectory of the MD simulation to be analyzed.
The commands needed to run GIST in cpptraj can be written in an input file (e.g gist.in) for convenience. The basic commands, assuming a given grid center and a grid size of $30\times30\times30$ voxels with \SI{0.5}{\angstrom} spacing, are:

@@

The first line specifies the amber topology file that was used to run MD. The second line species the trajectory file.
Note that the solute should not move significantly for a GIST calculation.
This is typically achieved by using restraints on the solute coordinates.
The third line tells \software{cpptraj} to run a GIST calculation using the region defined and print the output of the calculation in a file called \inlinecode{gist.dat}. The last instructs cpptraj to run the calculation.
The syntax for running \software{cpptraj} given this input file is:

@@

The \inlinecode{gist} action has a large number of options to modify the GIST calculation or its' output. A full list is provided at the end of this tutorial or through cpptrajs manual. The major options are summarized in the following table:

\begin{table}[h]
	\caption{The major options available in cpptrajs' GIST implementation.}\label{tab:GIST_options}
	\small
	\begin{tabularx}{\columnwidth}{@{}llXl@{}}
		\toprule
		Command       & Options & Explanation & Default  \\
		\midrule
		gridcntr    & <y> <y> <z> & Coordinates of the grid center in \AA & 0 0 0  \\
		griddim     & <nx> <ny> <nz> & Number of voxels along each axis defining the grid size & 40 40 40\\
		gridspacn & <dval> &Grid spacing / voxel dimensions in  \AA & 0.5    \\
		refdens         & <rdval> & Reference density of the bulk water in molecules/\AA$^3$ . & 0.0334    \\
		\bottomrule
	\end{tabularx}
\end{table}

Setting the reference density according to the water model used in your simulation is important to get correct results.
Densities of typical water models can be found in the \software{cpptraj} manual.
For other models, the density should be computed from a pure water simulation.

\paragraph{Postprocessing and Analyzing GIST results}
Cpptraj outputs the quantities calculated as a "open data explorer" (.dx) file format which can be easily visualized in Pymol or VMD.
Aside from the thermodynamic quantities, GIST also calculates the number densities of hydrogen and oxygen in each voxel. Here is a list of the most important quantities calculated by GIST:

\begin{itemize}
	\item{[gX] For every element in the main solvent, the number density of atoms found in the voxel, in units of the bulk density. If the same element occurs multiple times, the bulk density is scaled accordingly. For water, gO and gH are produced.}
	\item{[Esw] Mean solute-water interaction energy density.}
	\item{[Eww] Mean water-water interaction energy density.}
	\item{[PME] (only if PME was used) water PME energy density.}
	\item{[dTStrans] First order translational entropy density.}
	\item{[dTSorient] First order orientational entropy density .}
	\item{[dTSsix] First order entropy density for the six-dimensional combined orientation and translation space.}
	\item{[dipole] Magnitude of mean dipole moment (polarization).}
	\item{[dipolex] x-component of the mean water dipole moment density}
	\item{[dipoley] y-component of the mean water dipole moment density}
	\item{[dipolez] z-component of the mean water dipole moment density}
\end{itemize}

Several dx files are written by default, such as the energies, entropies, number density of oxygen, and number density of hydrogen. All calculated quantities are written to the output file (output.dat) and can be used to generate DX files for other quantities. A full list of all calculated properties found in the output in addition to the ones mentioned above is provided at the end of the tutorial.

\section{Tutorial}
Here, we aim to guide the reader through a GIST analysis of the streptavidin-biotin complex.
During the tutorial, we will
\begin{enumerate}
	\item set up and run GIST calculations for a small molecule as well as a protein binding pocket.
	\item compute the free energy of solvation for a small molecule using GIST.
	\item visualize local contributions to $\dgsolv$ in the streptavidin binding pocket as well as around biotin.
	\item compute the binding free energy of the complex from the individual energy contributions combined with the contribution of hydration computed using GIST.
	\item explore advantages and disadvantages of the GIST method when used to calculate  $\dgsolv$
\end{enumerate}
The tutorial is divided into a beginners part and an advanced part. In the beginners part, we will take a look at how to run GIST calculations for small molecules and interpret the output for a small molecule. In the advanced part, we will apply this knowledge to a protein binding pocket, visualize $\dgsolv$ hotspots and estimate a free energy of binding for a small molecule into this pocket using an end-state approach. In addition, while the beginners part focuses only on GIST calculation run, we will discuss all steps along the way in detail in the advanced section.

@@

\subsection{Tutorial data}
We will use the 1STP crystal structure to start our calculations. For the beginners section, we will use the biotin ligand as extracted from the crystal structure. For the second part, we will use the whole crystal structure.
For the sake of simplicity, you can download prepared and solvated Amber topologies and structures based on 1STP by running
\begin{lstlisting}[style=bash]
git clone git@github.com:liedllab/gist-tutorial.git
\end{lstlisting}
This will download the tutorial files (including both the manuscript and the examples in the \software{code} folder).
\
\subsection{Running GIST for Biotin}
In this example, we will be running a simple GIST calculation for a small molecule to get familiar with the usage of the GIST command in \software{cpptraj}.
\subsubsection{System preparation}
First, extract biotin from the crystal structure and generate a PDB file containing only biotin. Prepare your structure by adding hydrogens, making sure to choose the correct protonation state. Parametrize biotin with the forcefield of your choice. Solvate the parametrized biotin, making sur to choose a water box of large enough size so that the water on the box edges behaves bulk-like (a wall distance of at least 15\AA is recommended). A good starting point (though a bit dated) for the parametrization and solvation of small molecules is AMBER's tutorial 4b (\url{http://ambermd.org/tutorials/basic/tutorial4b/}). 
\subsubsection{Equilibration}
Equilibrate your solvated biotin to the conditions of your choice. Even when running a NVT simulation as the production run for the GIST analysis, it is advisible to include an NPT equilibration step to adjust the box size before hand. Allow enough equilibration time in the ensemble of your choice to allow the water to relax around your biotin structure.

\subsubsection{MD simulation}
\todo{Move simulation part from complex to here}

\subsection{Applying GIST for the Streptavidin-Biotin complex}
\subsubsection{System Preparation and Equilibration}
If you want to skip this section, run 
\begin{lstlisting}[style=bash]
make equilibration-targets
\end{lstlisting}
in the code folder. This assumes that you have Python and Amber set up properly and that \software{pmemd.cuda} is in the PATH.

@@

\section{Theory}
\label{sec:theory}
GIST is an implementation of inhomogeneous fluid solvation theory (IST) \cite{Lazaridis1998} that discretizes the free energy of solvation $\dgsolv$ on a three-dimensional grid. 
It was first devised by Nguyen et al. \cite{Nguyen2012} and its implementation in \software{cpptraj} was thoroughly described by Ramsey et al. \cite{Ramsey2016}. % to overcome the limitation of IST which was commonly used for regions of high water occupancy and its implementation in \software{cpptraj} was thoroughly described in \cite{Ramsey2016}.
Here, we only provide a short overview of the theory behind GIST.
For more detailed information, we recommend one of the more recent publications on developments in GIST \cite{Kraml2020,Chen2021}.

@@

\begin{checklist}{The CPPTRAJ GIST action}
